---
title: "01_datawork"
author: "Munkhtsetseg"
---


Library

```{r}
library(tidyverse)
library(ggplot2)
```

# Import the dataset and remove the duplicates

Import the dataset from the directory of: \~/Data Input/Preprocessing data/Preprocessing data.csv, assign the dataset as object of df:

```{r}
df <- read.csv("~/Data Input/Preprocessing data/Preprocessing data.csv")
```

Remove the duplicates with the function of distinct(), assign the dataset as df_01:

```{r}
df_01 <- df %>%
  distinct()
```

## Produce a table with missing data

For date options as year, month, etc:

```{r}
Table_NA_by_Year <- df_01 %>%
  rowwise %>%
  group_by(Station.name,Year) %>% 
  summarise(NA_date = sum(!is.na(Date)),
            NA_PM2 = sum(is.na(PM2)),
            NA_PM10 = sum(is.na(PM10)),
            NA_Vis = sum(is.na(Visibility)),
            NA_WD = sum(is.na(WD)),
            NA_WS = sum(is.na(WS)),
            NA_OPC = sum(is.na(OPC))
            ) 
Table_NA_by_Year

```

For station

```{r}
Table_NA_by_Station <- df_01 %>%
  rowwise %>%
  group_by(Station.name) %>% 
  summarise(NA_date = sum(!is.na(Date)),
            NA_PM2 = sum(is.na(PM2)),
            NA_PM10 = sum(is.na(PM10)),
            NA_Vis = sum(is.na(Visibility)),
            NA_WD = sum(is.na(WD)),
            NA_WS = sum(is.na(WS)),
            NA_OPC = sum(is.na(OPC))
  ) 
Table_NA_by_Station

```

By percentages

```{r df_01}
Table_NA_by_Station_percentage <- df_01 %>%
  rowwise %>%
  group_by(Station.name, Year) %>% 
  summarise(NA_date = sum(!is.na(Date)),
            NA_PM2 = sum(is.na(PM2)),
            NA_PM10 = sum(is.na(PM10)),
            NA_Vis = sum(is.na(Visibility)),
            NA_WD = sum(is.na(WD)),
            NA_WS = sum(is.na(WS)),
            NA_OPC = sum(is.na(OPC))
  ) %>% 
  mutate(
            NA_PM2p = if_else(NA_date != NA_PM2, NA_PM2/NA_date*100, 0)
  ) %>%
  do({q <- .$NA_PM2*100/.$NA_date; sdq <- sd(q); attr(sdq, "data.all") <- q; data.frame(sdq)}) 

 Table_NA_by_Station_percentage
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

# Remove the spikes, and produce an extended table

Remove the spikes in the datasets, and produce the table with NA, with removed spikes; express it in a percentages. 

### Remove the spikes Method 1. Mean value +- (3-5)SD

Method 2. Seasonal variations, and trend-mean

## Save dataset in folder: 01_data_raw

```{r}

```

# Tidy data

## Fill the missing data

Method 1. Fill the gap Method 2. Relationship equation Method 3. Look-up table

## Save dataset in folder: 02_data_tidy

```{r}

```
---
title: Explore Earthquakes
author: Charlotte Wickham
---

```{r}
library(tidyverse)
```


Read a clean version of data:

```{r}
la_palma <- read_csv("la-palma.csv")
```

Create spatial plot:

```{r}
#| label: fig-spatial-plot
#| fig-cap: "Locations of earthquakes on La Palma since 2017"
#| fig-alt: "A scatterplot of earthquake locations plotting latitude
#|   against longitude."
la_palma |> 
  ggplot(aes(Longitude, Latitude)) +
  geom_point(aes(color = Magnitude, size = 40-`Depth(km)`)) +
  scale_color_viridis_c(direction = -1) + 
  scale_size(range = c(0.5, 2), guide = "none") +
  theme_bw()
```

